Correlation among smart grid actors

### 2015-06-23 Article.tex
intro updates

Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences. One natural condition for stability revolves around sustaining a power generation that meets the demand at any time. If one is larger than the other, the system deviates from its synchronous state. If nothing is done to return to the synchronized equilibrium, this can lead to catastrophic cascades of failures. With traditional power plants based on fossil energies, the production can be scheduled in order to sustain the predicted consumption. Relatively small deviations from this schedule can then be supported almost on real time by using fast power plants, interconnections with border countries, as well as regulator entities. Even if most individual consumers may not realize it, the electicity prices are not constant rates and evolve with the production/consumption conditions on the grid. Operations are decided on a market where contracts are created based on information from the grid. Such a market environment appears necessary in order to improve the reliability of the whole system, and is already being used with large renewable plants in the production portfolio.

The smart grid vision however goes further and revolutionalizes this top-down centralized architecture by assuming that bidirectional power flows are allowed. This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution network. Nodes that were simply pure loads yesterday could behave tomorow as generators or loads depending on time. On the other hand, the penetration of renewables in the production is continuously growing, and is expected to become majority in a near future. These plants completely rely on the presence and the intensity of their respective ressources (wind, sun, tiles...), which predictions can be difficult and uncertain. Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition. The key to drive safely such a system is assumed to be information, and more precisely, the capacity to measure, communicate, and analyze data on real time.

@@

In this paper, we focus on these nodes in the distribution network that can produce and consume electricity. More precisely, we consider a set of agents that can be seen as aggregations of renewable distributed energy ressources (DER) and electrical loads. The production of an agent can be, of course, used to meet its own demand, but in cases where it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid. Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).

The constraints necessary for the electrical grid to remain in a stable state obiously forbid any anarchic system where prosumers inject power whenever and however they want. In this paper, the envisaged solution for organizing such a multi-producers system consists in a market environement, where participating entities announce an estimated production capacity for an upcoming period of time. If a contract is bought, the concerned entity commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.

In this paper, we consider that the production of a prosumer comes from renewables. The over-producing state of the agents, and thus the power they inject in the grid, might be therefore rather unstable. It has been suggested that energy diversification, geographical expansion, as well as using storage devices are envisageable techniques for  stabilizing the production. Besides, it has been argued that considering coalitions of prosumers as potential providers, and allowing these aggregations to enter the market as virtual power plants could have a stabilizing effect. Indeed, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio. One thus expect a more stable and predictable energy production for a coalition than for a single agent. However, all possible aggregations are not equal in terms of risks and expected productions. This is a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?

We will see that variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions. Understanding the correlation reliationships among the agents can thus be illuminating in the sense that it gives an indication about what coalitions to form and how much they should sell. More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different agents aggregations. We build a framework in which the system operator has the possibility of specifying both a maximum amount of risk and a minimum production acceptable to enter the market. Based on predictions, coalitions can decide wether or not to enter and what power quantities they are willing to provide. Of course, coalitions failing at fulfilling their obligations during the contracted period of time (because of bad forecasts, unexpected rare events, or intentional false information) will be exposed to financial penalities.

@@

\begin{figure}
\includegraphics[scale=.6]{real_utility.pdf}
\caption{Utility of random coalitions depending on their size. Lines shows mean utility values and colored areas stands for standard deviations. The slim blue noisy curve shows the real mean utility, and the thick red curve is its smoothed version by applying a Savitzky-Golay filter. On this plot the $ \alpha $ parameter of the utility function was selected according to equation \ref{eq:alpha_star} in order to favor 10 agents coalitions.}
\label{fig:real_utility}
\end{figure}

>>>>

Designing stable power systems is a classical engineering challenge since blackouts can have catastrophic consequences. One obvious condition for stability revolves around sustaining at any time a power generation that meets the demand. If one is larger than the other, the system deviates from its synchronous state. If nothing is done to return to the synchronized equilibrium, this can lead to catastrophic cascades of failures. With traditional power plants based on fossil energies, the production can be scheduled in order to sustain the predicted consumption. Deviations from this schedule can then be supported almost on real time by using fast response power plants, interconnections with border countries, as well as regulator entities. Even if most individual consumers may not realize it, the electicity prices are not constant rates and evolve with the production/consumption conditions on the grid. It is often assumed that the granularity of these prices is meant to increase in the future as a way to reverberate the production conditions on the end users. The pricing of electricity and the necessity for the grid operator to have different kinds of emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market. Such a market environment clearly necessitates communication in order to monitor and obtain information from the grid as well as exchange between participants. These kind of settings are already being used at the transport level with large renewable plants in the production portfolio, and they appear as a way to manage the reliability of the whole system, especially when the penetration of renewables is important.

The smart grid vision however goes further and revolutionalizes this top-down centralized architecture by assuming that bidirectional power flows are allowed. This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution networks. Nodes that were simply pure loads yesterday could behave tomorow as generators or loads depending on time. On the other hand, the penetration of renewables in the production is continuously growing, and is expected to become majority in a near future. These plants completely rely on the presence and the intensity of their respective ressources (wind, sun, tiles...), which predictions can be difficult and uncertain. Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition. The key to drive safely such a system is assumed to be information, and more precisely, the capacity to measure, communicate, and analyze data on real time.

@@

In this paper, we focus on these "\textit{nodes}" in the distribution network that can produce and consume electricity. More precisely, we consider a set of agents that can be seen as aggregators of renewable distributed energy ressources (DER) and electrical loads. The production of an agent can be, of course, used to meet its own demand, but in cases where it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid. Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).

As beautiful as this distributed participatory production scenario seems to be, the constraints necessary for the electrical grid to remain in a stable state obiously forbid any anarchic system where prosumers inject power whenever and however they want. Actually the architecture and control of such a distributed grid creates numerous interdisciplinary challenges. In this paper, the envisaged solution for organizing such a multi-producers system consists in extending the kind of market environement discussed in the previous section. Participating aggregators estimate a production capacity for an upcoming period of time and announce it on the market. If a contract between the grid and a participant is bought, the latter commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.

In this paper, we consider that the production of a prosumer comes from renewables. Contracts are therefore based on predictions and come with a given amount of risk due to the intermittent nature of most renewables. Using storage as buffers in order to reduce riks is a popular approach. Actually, a whole branch of the smart grid litterature is even studying the possibility to use electric vehicles as moving storage capacities for stabilizing the grid. Without storage and proper control, the over-producing state of the agents, and thus the power they inject in the grid, might be rather unstable. This is clearly unacceptable for the grid operator that cannot ensure system stability if it has to deal with numerous small unpredictable entities.

Forming coalitions/aggregations is an envisaged solution to both the number of entities sollicitating the operator, and their high variabilities. Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side, since an aggregator can be considered as a single node by the level above it. On the other hand, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio. One thus expect a more stable and predictable energy production for a coalition than for a single agent. Nevertheless, all coalitions are obviously not equally stable, which means that special attention should be paid to the aggregation step.

Furthermore, in our settings, since the prosumers have both a consumption and a production component, each depending on location and time, the diversification of the aggregations seems not trivial. This is a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?

We will see that variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions. Understanding the correlation reliationships among the agents can thus be illuminating in the sense that it gives an indication about what coalitions to form and how much they should sell. More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions. We build a framework in which the system operator has the possibility of specifying both a maximum amount of risk and a minimum production acceptable to enter the market. Based on predictions, coalitions can decide wether or not to enter and what power quantities they are willing to provide. Of course, coalitions failing at fulfilling their obligations during the contracted period of time (because of bad forecasts, unexpected rare events, or intentional false information) will be exposed to financial penalities.

@@
